TacticToe: Learning the Game of Tactics

### 2018-02-07 tactictoe_jv.tex
future work: wip

\usepackage[final]{listings}
%\usepackage[scaled=.95]{couriers}
\lstdefinelanguage{SML}%
{columns=fullflexible,
  keywords={THEN,THENL,val,open,store_thm,fun,fn},%
  frame=none,
  sensitive=true,
  keywordstyle=\fontfamily{lmss}\footnotesize\selectfont,%
  basicstyle=\fontfamily{pcr}\selectfont,%
  stringstyle=\tt,
  morestring=[b]",
  literate={\ }{$\mkern6mu$}1%
   {=}{{\tt\raisebox{-.15mm}{=}}}1%
   {[}{{\tt\raisebox{-.15mm}{[}}}1%
   {]}{{\tt\raisebox{-.15mm}{]}}}1%
   {->}{{$\rightarrow$}}1%
   {==>}{{$\Rightarrow$}}1%
   {wedge}{{$\wedge$}}1%
   {>=}{{$\geq$}}1%
   {'a}{{$\alpha$}}1%
   {'b}{{$\beta$}}1%
   {ldots}{$\ldots$}1%
   {!}{$\forall$}1%
   {``}{\hspace{-.5mm}`\hspace{-1mm}`}1%
  }

\lstdefinelanguage{SMLSmall}%
{columns=fullflexible,
  keywords={THEN,THENL,val,open,store_thm,fun,fn},%
  frame=none,
  sensitive=true,
  keywordstyle=\fontfamily{lmss}\scriptsize\selectfont,%
  basicstyle=\fontfamily{pcr}\small\selectfont,%
  stringstyle=\tt,
  morestring=[b]",
  literate={\ }{$\mkern6mu$}1%
   {=}{{\tt\raisebox{-.15mm}{=}}}1%
   {[}{{\tt\raisebox{-.15mm}{[}}}1%
   {]}{{\tt\raisebox{-.15mm}{]}}}1%
   {->}{{$\rightarrow$}}1%
   {==>}{{$\Rightarrow$}}1%
   {wedge}{{$\wedge$}}1%
   {<=}{{$\leq$}}1%
   {>=}{{$\geq$}}1%
   {'a}{{$\alpha$}}1%
   {'b}{{$\beta$}}1%
   {ldots}{$\ldots$}1%
   {!}{$\forall$}1%
   {fff}{{DB.fetch}}1%
   {``}{\hspace{-.5mm}`\hspace{-1mm}`}1%
  }
\lstset{language=SML}
\begin{document}
\maketitle

@@

\begin{lstlisting}[language=SMLSmall]
tactictoe ([], ``countable (UNIV:num set)``);
let fun RWTAC thms = SRW_TAC [] thms in RWTAC [(fetch "pred_set"
"countable_def"), (fetch "pred_set" "INJ_DEF")] THEN Q.EXISTS_TAC `\\x.x` THEN
METIS_TAC [] end
\end{lstlisting}

\begin{lstlisting}[language=SMLSmall]
tactictoe ([], ``count (n+m) DIFF count n = IMAGE ($+n) (count m)``);
SRW_TAC [ARITH_ss] [(fetch "pred_set" "EXTENSION"), EQ_IMP_THM] THEN
Q.EXISTS_TAC `x - n` THEN SRW_TAC [ARITH_ss] []
\end{lstlisting}

@@

There is plenty of future work in the directions opened here.
To improve the quality of the predicted tactics,
we would like to predict their arguments independently.
To be even more precise, the relation between the
tactic arguments and their respective goals could be used.
Additionally, we could aim for a tighter combination with the ATP-based hammer
systems. This would perhaps make \tactictoe slower, but it might allow
finding proofs that are so far both beyond the ATPs and \tactictoe's
powers. The idea of reusing high-level blocks of reasoning and
then learning their selection could also be explored
further in various contexts. Larger frequent blocks of (instantiated) tactics
in ITPs as well as blocks of inference patterns in ATPs could be detected
automatically, their
usefulness in particular proof situations learned from the large corpora of
ITP and ATP proofs, and reused in high-level proof search.

>>>>

\usepackage[final]{listings}
%\usepackage[scaled=.95]{couriers}
\lstdefinelanguage{SML}%
{columns=fullflexible,
  keywords={THEN,THENL,val,open,store_thm,fun,fn,let,in,end},%
  frame=none,
  sensitive=true,
  keywordstyle=\fontfamily{lmss}\footnotesize\selectfont,%
  basicstyle=\fontfamily{pcr}\selectfont,%
  stringstyle=\tt,
  morestring=[b]",
  literate={\ }{$\mkern6mu$}1%
   {=}{{\tt\raisebox{-.15mm}{=}}}1%
   {[}{{\tt\raisebox{-.15mm}{[}}}1%
   {]}{{\tt\raisebox{-.15mm}{]}}}1%
   {->}{{$\rightarrow$}}1%
   {==>}{{$\Rightarrow$}}1%
   {wedge}{{$\wedge$}}1%
   {>=}{{$\geq$}}1%
   {'a}{{$\alpha$}}1%
   {'b}{{$\beta$}}1%
   {ldots}{$\ldots$}1%
   {!}{$\forall$}1%
   {``}{\hspace{-.5mm}`\hspace{-1mm}`}1%
  }

\lstdefinelanguage{SMLSmall}%
{columns=fullflexible,
  keywords={THEN,THENL,val,open,store_thm,fun,fn,let,in,end},%
  frame=none,
  sensitive=true,
  keywordstyle=\fontfamily{lmss}\scriptsize\selectfont,%
  basicstyle=\fontfamily{pcr}\small\selectfont,%
  stringstyle=\tt,
  morestring=[b]",
  literate={\ }{$\mkern6mu$}1%
   {=}{{\tt\raisebox{-.15mm}{=}}}1%
   {[}{{\tt\raisebox{-.15mm}{[}}}1%
   {]}{{\tt\raisebox{-.15mm}{]}}}1%
   {->}{{$\rightarrow$}}1%
   {==>}{{$\Rightarrow$}}1%
   {wedge}{{$\wedge$}}1%
   {<=}{{$\leq$}}1%
   {>=}{{$\geq$}}1%
   {'a}{{$\alpha$}}1%
   {'b}{{$\beta$}}1%
   {ldots}{$\ldots$}1%
   {!}{$\forall$}1%
   {fff}{{DB.fetch}}1%
   {``}{\hspace{-.5mm}`\hspace{-1mm}`}1%
  }
\lstset{language=SML}
\begin{document}
\maketitle

@@

\begin{lstlisting}[language=SMLSmall]
tactictoe ([], ``countable (UNIV:num set)``);

let fun RWTAC thms = SRW_TAC [] thms in 
  RWTAC [fetch "pred_set""countable_def", fetch "pred_set" "INJ_DEF"] THEN 
  Q.EXISTS_TAC `\\x.x` THEN
  METIS_TAC []
end
\end{lstlisting}

\begin{lstlisting}[language=SMLSmall]
tactictoe ([], ``count (n+m) DIFF count n = IMAGE ($+n) (count m)``);

SRW_TAC [ARITH_ss] [fetch "pred_set" "EXTENSION", EQ_IMP_THM] THEN
Q.EXISTS_TAC `x - n` THEN 
SRW_TAC [ARITH_ss] []
\end{lstlisting}

@@

There is plenty of future work in the directions opened here.

To improve the quality of the predicted tactics,
we would like to predict other type of arguments independently, as it was done 
for theorems. In this direction, the most interesting type of arguments to 
predict next are terms as they are ubiquitous in tactics. Further along the way,
we would like to be to able to create new tactics not only by changing their 
arguments but also by programming it with any construction available in a
programming language such as \sml.

There is still something to be gain by having a better guidance for the search. 
That is why se also want to experiment with stronger machine learning algorithm 
such as deep neural networks as a model for the policy and evaluation in MCTS.
This quality of this model could be enhanced by additional data augmentation by 
training and testing existing and synthesized tactics.
Another possible way to guide the proof search is by conjecturing intermediate 
steps \cite{}, allowing to prove a difficult theorem by decomposing it into 
multiple easy steps.
